\phantomsection
\chapter{Feature extraction}
\label{chap:extraction}

\noindent After having detected the face, for example using Viola-Jones algorithm, it is necessary to perform feature extraction in order to process data before classification. 
This step takes an image as input, and extracts vectors characterizing its main features. In the case of a facial expression recognition system, resulting vectors contain informations about spacial configuration of facial features, but can also encode informations about shape, texture or movement of the image's content \cite{CHI03}.
\newline

\noindent There are however many kinds of algorithms outputting features vectors, and the choice of an effective one depends on many criteria. These feature extration methods can generally be ranked among 2 categories: they can either be appearance-based or geometry-based, depending on the way they extract feature vectors. The aim of this chapter is to explain the differences between these 2 types of feature extraction methods, and provide some examples from each category.
\newline

\phantomsection
\section{Appearance-based methods}
% PCA, LDA, Gabor wavelets, Eigenfaces

\noindent As said in the introduction, the Appearance-based method aims to extract the appearance changes of the face. In order to do that the method uses image filters that are applied to the whole face or that are applied to specific parts of the face \cite{SHA09}. With this kind of method, if there is any change in the lightning or the pose of the head, the recognition will be less effective.
\newline

\noindent The most used Appearance-based method is the Gabor wavelets because of their performance. The disadvantage of the Gabor wavelets method is that computing the whole face with the bank of gabor filters is time and memory consuming \cite{SHA09}.
\newline

\phantomsection
\section{Geometry-based methods}
% LBP

\noindent As said in the introduction, the Geometry-based method aims to locate important facial components. The shape and the location of these facial features are extracted from an face image and then are concatenated into a features vector. This features vector represents the geometry of the face \cite{SHA09}.
\newline

\noindent Compared to Appearance-based method, the Geometry-based method has equal or better results. This was recently tested by Valstar et al. \cite{VAL05} \cite{VAL06}.
\newline

\noindent The disadvantage of this method is that it needs reliable and accurate feature detection and also tracking if necessary. And this can involves a lot of extra processing and so can be time consuming \cite{SHA09}.
\newline

\noindent Conclusion paragraph that opens on LBP

