%\phantomsection
\thispagestyle{plain}
\hypersetup{bookmarksdepth=-2} % so it doesn't appear on the toc nor in the pdf bookmarks
\addcontentsline{toc}{chapter}{Preface}
\chapter*{Preface}
\hypersetup{bookmarksdepth}%back to tocdepth

\noindent This report documents the semester project entitled \textit{Facial expression recognition using Local Binary Patterns}. The project was carried out during the 9th semester of specialization \textit{Vision, Graphics, and Interactive Systems} under the Department of Electronic Systems at Aalborg University in Autumn 2012. 
\newline

\noindent The report is divided into four parts plus appendices: \textit{Introduction}, \textit{Feature Detection}, \textit{Feature Classification}, \textit{Implementation} and \textit{Evaluation}. The first part review the general structure of a facial expression recognition system and its main issues, and concludes with a state of the art of existing systems. Analysis of possible solutions and design of our system are contained in the following two parts, and the fourth part describes our implementation. The last part evaluates the performance and accuracy of our system and concludes on the project as a whole. 
\newline

\noindent References to secondary literature sources are made using the syntax [number]. The number refers to the alphabetically sorted bibliography found at the end of the report, just before the appendices.
\newline

\noindent We would like to thank our supervisor at Aalborg University Zheng-Hua Tan for supporting us in this challenging project. 
\newline

\noindent A CD is attached to this report which includes:
\begin{itemize}
\item Source code of the developed program.
\item PDF file of this report.
\end{itemize}

\vspace{\baselineskip}\hfill Aalborg University, \today
\vfill\noindent
\begin{minipage}[b]{0.45\textwidth}
 \centering
 \rule{\textwidth}{0.5pt}\\
  Maxime Coupez\\
 {\footnotesize <mcoupe12@es.aau.dk>}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\textwidth}
 \centering
 \rule{\textwidth}{0.5pt}\\
  Kim-Adeline Miguel\\
 {\footnotesize <kmigue12@es.aau.dk>}
\end{minipage}
\vspace{3\baselineskip}
\begin{center}
\begin{minipage}[b]{0.45\textwidth}
 \centering
 \rule{\textwidth}{0.5pt}
  Julia Alexandra Vigo\\
 {\footnotesize <jvigo12@es.aau.dk>}
\end{minipage}
\end{center}
